---
title: "Simulation of imputation of censored values for June 30th Zoom Call v3"
author: "Marc Roddis"
date: "6/29/2020"
output: html_document
---

```{r chunk1, include=FALSE}
library(RCurl)
library(tidyverse)
library(styler)
library(gridExtra)
library(cowplot)
library("NADA")
library(truncreg)
library(truncnorm)
knitr::opts_chunk$set(echo=FALSE)
```

### Finding appropriate simulation parameters from observed data

We created the test dataset `testdata_cen_omit` from the original observed data `pcb.csv` by omitting all missing values of `CB28` and `CB153`, removing all observations except those from herring species, removing all observations prior to 1989, re-indexing 1989 as "year zero", removing all variables except `YEAR`, `CB28` and `CB153`, omitting all censored observations, and replacing concentrations with log-concentrations.

```{r chunk2, include=FALSE}
pcb_df <- read_csv("pcb.csv")
pcb_tib <- as_tibble(pcb_df)
testdata1 <- pcb_tib %>%
  mutate(CB28 = ifelse(CB28< -8, NA, CB28) ) %>%
  mutate(CB153 = ifelse(CB153< -8, NA, CB153) ) %>%
  mutate(CB28 = ifelse(CB28> -0.0001 & CB28< 0.0001, NA, CB28) ) %>%
  filter(!is.na(CB28)) %>%
  filter(!is.na(CB153)) %>%
  filter(SPECIES == "Herring") %>%
  filter(YEAR >= 1989) %>%
  mutate(YEAR = YEAR-1989) %>%
  select(YEAR, CB28, CB153)
```

```{r chunk3, include=FALSE}
geomean <- function(x) round(exp(mean( log(x) ) ),4)
testdata_cen_omit <- testdata1 %>%
  filter(CB28>0) %>%
  filter(CB153>0) %>%
  group_by(YEAR) %>%
  summarise_at(vars(CB28,CB153), geomean  ) %>%
  ungroup()
testdata_cen_omit_logconc <- testdata_cen_omit %>%
  mutate(CB28 = log(CB28)) %>%
  mutate(CB153 = log(CB153))
```

Fitting linear models to the test data gave the following fixed parameters: $$CB28 = -4.54 - 0.046*YEAR$$ $$CB153 = -1.85 - 0.039*YEAR$$ $$CB28 = -3.12 + 0.86*CB153$$ $$sd(CB28)=0.46$$ $$sd(CB153)=0.46$$


```{r chunk4}
# attach(testdata_cen_omit_logconc) # assume logconcs have normal distribution
# summary(lm(CB28~YEAR)) # -4.54 - 0.046*YEAR
# summary(lm(CB153~YEAR)) # -1.85 - 0.039*YEAR
# summary(lm(CB28~CB153)) # -3.12 + 0.86*CB153
# summary(lm(CB28~CB153+YEAR)) #  -3.60 + 0.51*CB153 - 0.026*YEAR
# sd(CB28) # 0.46
# sd(CB153) # 0.46
# quant_test <- quantile(testdata_cen_omit_logconc$CB28, probs=0.1, names=FALSE ) # (-5.76, -5.32, -5.22, -4.79)
# quant_test
```

#### Simulation and model fitting assuming no censored values

These fixed parameters were used to generate the dataset `data_geomean2` as follows:

12 values for the log-concentration of CB153 per year, for ten years, were generated and denoted as `cb153sim`.

From every such CB153 value, the corresponding value for CB28 was generated.

Annual geometric means for CB28 and CB153 concentrations were generated. 


We will first fit a linear model to `data_geomean2`, assuming no censored values; this model `best_fit` is the gold standard against which will will later evaluate our three methods for dealing with censored values: omission, substitution, imputation from cenreg().  We will also test each method at various LOQ values:  For each method, at each LOQ value, we will fit the corresponding linear model and compute the mse, squared-bias, and variance, for each year separately.   We will then use these results to evaluate these three methods.

The code chunk below generates `data_geomean2` for 200 iterations, fits a linear model `best_fit` at each iteration, and computes the corresponding mse, squared-bias and variance.

```{r chunk5, include=FALSE}
cprop <- 0.5 # censoring proportion
n_iter <- 200
set.seed(1)
years <- c(0:9)
geomean_prelogged <- function(x) exp(mean(x, na.rm = TRUE))

get_sim_data_best = function(sample_size=12) { 
  noise153 = rnorm(n = sample_size*10, mean = 0, sd = 0.5) # sd is from obs data
  noise28 = rnorm(n = sample_size*10, mean = 0, sd = 0.5) # sd is from obs data
  year = rep(years, sample_size)
  cb153 = -1.85 - 0.039 * year + noise153
  cb28 = -3.12 + 0.86*cb153 + noise28
  df = data.frame(year,cb28,cb153)
  data_tib2 = as_tibble(df)
  data_geomean2 <- data_tib2 %>%
    group_by(year) %>%
    summarise_at(vars(cb28, cb153), geomean_prelogged  ) %>%
    ungroup()
  data_geomean2
}

estimates_best <- matrix(0, nrow = 10, ncol = n_iter)

for (iter in 1:n_iter) {

  # simulate new, random, training data
  # this is the only random portion of the bias, var, and mse calculations
  # this allows us to calculate the expectation over the data

  sim_data_best <- get_sim_data_best(sample_size=12)
  best_fit = lm(sim_data_best$cb28 ~ sim_data_best$year)
  estimates_best[, iter] = predict(best_fit)
  data.frame(estimates_best)
}

best_annual_concs <-  exp(-4.711 - 0.03354 * years)
best_annual_means_n_iter <- rep(best_annual_concs, n_iter)
best_annual_means_matrix <- matrix(best_annual_means_n_iter, nrow=10, ncol=n_iter)
best_residuals_df <- data.frame(estimates_best - best_annual_means_matrix)

g = function(x) {
  mean(x^2)
}

best_mse_vector <- apply(best_residuals_df, 1, g)
best_mse_df <- data.frame(best_mse_vector)
estimates_best_mean = apply(estimates_best, 1, mean)
best_bias_sq_vector = (estimates_best_mean - best_annual_concs)^2
best_bias_sq_df = data.frame((best_bias_sq_vector))
best_variance_vector = apply(estimates_best, 1, var)
best_variance_df = data.frame(best_variance_vector)
best_mse_bias_sq_var_matrix = matrix(0, nrow = 10, ncol=3)
best_mse_bias_sq_var_matrix[,1] <- best_mse_vector
best_mse_bias_sq_var_matrix[,2] <- best_bias_sq_vector
best_mse_bias_sq_var_matrix[,3] <- best_variance_vector
best_mse_bias_sq_var_df <- data.frame(best_mse_bias_sq_var_matrix)
colnames(best_mse_bias_sq_var_df) <- c("mse_best","bias_best","variance_best")
best_mse_bias_sq_var_df

```

#### Simulation and model fitting after omitting censored values

We now generate the dataset `data_omit_geomean` from the fixed parameters as follows:

12 values for the log-concentration of CB153 per year, for ten years, were generated and denoted as `cb153sim`.

From every such CB153 value, the corresponding value for CB28 was generated.

`median(cb28)` was used as the level of quantification `LOQ_p50`.

Observations with CB28 < LOQ_p50 were removed from the dataset.

Annual geometric means for CB28 and CB153 concentrations were generated. 

The code chunk below generates `data_omit_geomean` for 1000 iterations, fits a linear model `omit_fit` at each iteration, and computes the corresponding mse, squared-bias and variance.

```{r chunk6, include=FALSE}

get_sim_data = function(sample_size=12) { 
  noise153 = rnorm(n = sample_size*10, mean = 0, sd = 0.5) # sd is from obs data
  noise28 = rnorm(n = sample_size*10, mean = 0, sd = 0.5) # sd is from obs data
  year = rep(years, sample_size)
  cb153 = -1.85 -0.039 * year + noise153 #beta_obs = -0.039
  # changing beta_obs value, throws an error 
  cb28 = -3.12 + 0.86*cb153 + noise28
  data.frame(year,cb28,cb153)
}

estimates_omit <- matrix(0, nrow = 10, ncol = n_iter)

for (iter in 1:n_iter) {
  sim_omit_data <- get_sim_data(sample_size=12)
  data_omit_tib <- as_tibble(sim_omit_data)
  data_omit_geomean <- data_omit_tib %>%
    filter(cb28 >= quantile(cb28,probs = 0.5, names=FALSE) ) %>%
    group_by(year) %>%
    summarise_at(vars(cb28, cb153), geomean_prelogged  ) %>%
    ungroup()
  omit_fit = lm(data_omit_geomean$cb28 ~ data_omit_geomean$year)
  estimates_omit[, iter] = predict(omit_fit)
  data.frame(estimates_omit)
}


best_annual_concs <-  exp(-4.711 - 0.03354 * years)
best_annual_means_n_iter <- rep(best_annual_concs, n_iter)
best_annual_means_matrix <- matrix(best_annual_means_n_iter, nrow=10, ncol=n_iter)

omit_residuals_df <- data.frame(estimates_omit - best_annual_means_matrix)

g = function(x) {
  mean(x^2)
}

omit_mse_vector <- apply(omit_residuals_df, 1, g)
omit_mse_df <- data.frame(omit_mse_vector)
estimates_omit_mean = apply(estimates_omit, 1, mean)
omit_bias_sq_vector = (estimates_omit_mean - best_annual_concs)^2

omit_bias_sq_df = data.frame((omit_bias_sq_vector))
omit_variance_vector = apply(estimates_omit, 1, var)
omit_variance_df = data.frame(omit_variance_vector)
omit_mse_bias_sq_var_matrix = matrix(0, nrow = 10, ncol=3)
omit_mse_bias_sq_var_matrix[,1] <- omit_mse_vector
omit_mse_bias_sq_var_matrix[,2] <- omit_bias_sq_vector
omit_mse_bias_sq_var_matrix[,3] <- omit_variance_vector
omit_mse_bias_sq_var_df <- data.frame(omit_mse_bias_sq_var_matrix)
colnames(omit_mse_bias_sq_var_df) <- c("mse_omit","bias_omit","variance_omit")
omit_mse_bias_sq_var_df
```

```{r chunk6b, include=FALSE}
estimates_subst <- matrix(0, nrow = 10, ncol = n_iter)

for (iter in 1:n_iter) {
  sim_subst_data <- get_sim_data(sample_size=12)
  data_subst_tib <- as_tibble(sim_subst_data)
  data_subst_geomean <- data_subst_tib %>%
    mutate(cb28 = ifelse(cb28<quantile(cb28,probs = cprop, names=FALSE), quantile(cb28,probs = cprop, names=FALSE) - log(sqrt(2) ), cb28)) %>%
    group_by(year) %>%
    summarise_at(vars(cb28, cb153), geomean_prelogged  ) %>%
    ungroup()
  subst_fit = lm(data_subst_geomean$cb28 ~ data_subst_geomean$year)
  estimates_subst[, iter] = predict(subst_fit)
  data.frame(estimates_subst)
}

subst_residuals_df <- data.frame(estimates_subst - best_annual_means_matrix)

subst_mse_vector <- apply(subst_residuals_df, 1, g)
subst_mse_df <- data.frame(subst_mse_vector)
estimates_subst_mean = apply(estimates_subst, 1, mean)
subst_bias_sq_vector = (estimates_subst_mean - best_annual_concs)^2

subst_bias_sq_df = data.frame((subst_bias_sq_vector))
subst_variance_vector = apply(estimates_subst, 1, var)
subst_variance_df = data.frame(subst_variance_vector)
subst_mse_bias_sq_var_matrix = matrix(0, nrow = 10, ncol=3)
subst_mse_bias_sq_var_matrix[,1] <- subst_mse_vector
subst_mse_bias_sq_var_matrix[,2] <- subst_bias_sq_vector
subst_mse_bias_sq_var_matrix[,3] <- subst_variance_vector
subst_mse_bias_sq_var_df <- data.frame(subst_mse_bias_sq_var_matrix)
colnames(subst_mse_bias_sq_var_df) <- c("mse_subst","bias_subst","variance_subst")
subst_mse_bias_sq_var_df
```

```{r chunk6c7, include=FALSE}

estimates_cenreg <- matrix(0, nrow = 10, ncol = n_iter)

for (iter in 1:n_iter) {
  dat <- get_sim_data(sample_size=12)
  dat_tib = as_tibble(dat)
  data_indic <- dat_tib %>%     # Martin's suggestion (must censor values)
    mutate(cb28 = pmax(cb28, quantile(cb28,probs = cprop, names=FALSE) )) %>%
    mutate(ci28 = cb28 < quantile(cb28,probs = cprop, names=FALSE) ) 

  cenreg_fit = cenreg(Cen(data_indic$cb28, data_indic$ci28) ~ data_indic$cb153, dist="gaussian")
  cenreg_alpha <- coef(cenreg_fit)[1]
  cenreg_beta <- coef(cenreg_fit)[2]
  cenreg_residual <- residuals(cenreg_fit)
  
#   truncreg(formula, data, subset, weights, na.action,
# point = 0, direction = "left",
# model = TRUE, y = FALSE, x = FALSE, scaled = FALSE, ...)
  
  cenreg_imputed <- dat %>%
    mutate(cb28 = ifelse(cb28<quantile(cb28,probs = cprop, names=FALSE), etruncnorm(a=-Inf, b=quantile(cb28,probs = cprop, names = FALSE), mean=cenreg_alpha + cenreg_beta * cb153, sd=cenreg_residual) , cb28) ) %>%
    group_by(year) %>%
    summarise_at(vars(cb28, cb153), geomean_prelogged  ) %>%
    ungroup()
  
  cenreg_fit_lm = lm(cenreg_imputed$cb28 ~ cenreg_imputed$year)
  
  estimates_cenreg[, iter] = predict(cenreg_fit_lm)
  data.frame(estimates_cenreg)
}

cenreg_residuals_df <- data.frame(estimates_cenreg - best_annual_means_matrix)

cenreg_mse_vector <- apply(cenreg_residuals_df, 1, g)
cenreg_mse_df <- data.frame(cenreg_mse_vector)
estimates_cenreg_mean = apply(estimates_cenreg, 1, mean)
cenreg_bias_sq_vector = (estimates_cenreg_mean - best_annual_concs)^2

cenreg_bias_sq_df = data.frame((cenreg_bias_sq_vector))
cenreg_variance_vector = apply(estimates_cenreg, 1, var)
cenreg_variance_df = data.frame(cenreg_variance_vector)
cenreg_mse_bias_sq_var_matrix = matrix(0, nrow = 10, ncol=3)
cenreg_mse_bias_sq_var_matrix[,1] <- cenreg_mse_vector
cenreg_mse_bias_sq_var_matrix[,2] <- cenreg_bias_sq_vector
cenreg_mse_bias_sq_var_matrix[,3] <- cenreg_variance_vector
cenreg_mse_bias_sq_var_df <- data.frame(cenreg_mse_bias_sq_var_matrix)
colnames(cenreg_mse_bias_sq_var_df) <- c("mse_cenreg","bias_cenreg","variance_cenreg")

```


### Results

The three graphs A, B, C below show the variation of MSE, squared-bias, and variance respectively, over the simulated 10-year period.  The red line corrresponds to model `omit_fit` (recall: omission of censored observations) and the black line corresponds to `best_fit`.

```{r chunk7, include=FALSE}

all_results_df <- data.frame(cbind(best_mse_bias_sq_var_df,omit_mse_bias_sq_var_df, subst_mse_bias_sq_var_df, cenreg_mse_bias_sq_var_df)) 
all_results_tib <- as_tibble(all_results_df)
all_results_tib <- all_results_tib %>%
  mutate(year=c(0:9)) %>%
  mutate(bias_sq_plus_variance_best = bias_best+variance_best) %>%
  mutate(bias_sq_plus_variance_omit = bias_omit+variance_omit) %>%
  mutate(bias_sq_plus_variance_subst = bias_omit+variance_subst) %>%
  mutate(bias_sq_plus_variance_cenreg = bias_omit+variance_cenreg) 

all_results_tib

plot_all_mse <- ggplot(data = all_results_tib) +
  # geom_line(mapping = aes(x = year, y = mse_best, color="best")) +
  # geom_line(mapping = aes(x = year, y = mse_omit, color="omit")) +
  geom_line(mapping = aes(x = year, y = mse_subst, color="subst")) +
  geom_line(mapping = aes(x = year, y = mse_cenreg, color="cenreg")) 

plot_all_bias <- ggplot(data = all_results_tib) +
  # geom_line(mapping = aes(x = year, y = bias_best, color="best")) +
  # geom_line(mapping = aes(x = year, y = bias_omit, color="omit")) +
  geom_line(mapping = aes(x = year, y = bias_subst, color="subst")) +
  geom_line(mapping = aes(x = year, y = bias_cenreg, color="cenreg"))

# bias_omit is so high that it is omitted from the following plot
plot_best_subst_cenreg_bias <- ggplot(data = all_results_tib) +
  # geom_line(mapping = aes(x = year, y = bias_best, color="best")) +
  geom_line(mapping = aes(x = year, y = bias_subst, color="subst")) +
  geom_line(mapping = aes(x = year, y = bias_cenreg, color="cenreg"))

plot_all_variance <- ggplot(data = all_results_tib) +
  # geom_line(mapping = aes(x = year, y = variance_best, color="best")) +
  # geom_line(mapping = aes(x = year, y = variance_omit, color="omit")) +
  geom_line(mapping = aes(x = year, y = variance_subst, color="subst")) +
  geom_line(mapping = aes(x = year, y = variance_cenreg, color="cenreg"))

# best_verify_bias_variance_decomp <- ggplot(data = all_results_tib) +
#   geom_line(mapping = aes(x = year, y = mse_best, color="mse_best")) +
#   geom_line(mapping = aes(x = year, y = bias_sq_plus_variance_best ))
# 
# omit_verify_bias_variance_decomp <- ggplot(data = all_results_tib) +
#   geom_line(mapping = aes(x = year, y = mse_omit, color="mse_omit")) +
#   geom_line(mapping = aes(x = year, y = bias_sq_plus_variance_omit ))
```

```{r chunk8}
# quantile(dat$cb28,probs = cprop, names = FALSE)
data_subst_geomean2 <- data_subst_tib %>%
  mutate(cb28 = ifelse(cb28<quantile(cb28,probs = cprop, names=FALSE), quantile(cb28,probs = cprop, names=FALSE) - log(sqrt(2) ), cb28)) %>%
  group_by(year) %>%
  summarise_at(vars(cb28, cb153), geomean_prelogged  ) %>%
  ungroup()

data_indic2 <- dat_tib %>%     # Martin's suggestion (must censor values)
    mutate(cb28 = pmax(cb28, quantile(cb28,probs = cprop, names=FALSE) )) %>%
    mutate(ci28 = cb28 < quantile(cb28,probs = cprop, names=FALSE) )
data_subst_geomean2$cb28
mean(data_subst_tib$cb28)
sum(data_indic2$ci28)
plot_grid(plot_all_mse, plot_all_bias, plot_best_subst_cenreg_bias, plot_all_variance, labels = "AUTO")
```

The two graphs A, B below show the variation of MSE (red curve) and squared-bias-plus-variance (black curve) from `best_fit` and `omit_fit` respectively, over the simulated 10-year period.   The famous result "Bias-variance decomposition" states $$ MSE = Bias^2 + Variance$$ so we expect the black and red curves to coincide (be superposed); happily they are :)

```{r chunk9}

# plot_grid( best_verify_bias_variance_decomp, omit_verify_bias_variance_decomp, labels = "AUTO")

```



